### 2014-09-26 metaTracts_supplymental_pvis2014.tex
supplemental folders

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{pdfpages}
\begin{document}
\input {./supplemental/supplementaryMaterial.tex}
\end{document}