### 2014-09-26 supplemental/supplementaryMaterial.tex

\section{Supplementary Material}
\label{sec:sm}
Section~\ref{subsec:usereval} contains the our user evaluation questionnaire.
\subsection{User Evaluation Questionnaire}
The questionnaire contains links to videos which were seen by the participants.
\label{subsec:usereval}
\includepdf[pages={1-13}]{./supplemental/MetaTracts_usereval.pdf}