### 2014-12-11 metaTracts_supplymental_pvis2014.tex

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{pdfpages}
\begin{document}
	\title{MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites \\
		Supplementary Material}
	\author{Arindam Bhattacharya\thanks{e-mail: arindamb86@gmail.com}\\ %
		\scriptsize The Ohio State University %
		\and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
		\scriptsize University of Applied Sciences - Upper Austria%
		\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
		\scriptsize The Ohio State University
		\and Artem Amirkhanov\thanks{e-mail:artem.amirkhanov@fh-wels.at}\\ %
		\scriptsize University of Applied Sciences - Upper Austria%
		\and Johann Kastner\thanks{e-mail:johann.kastner@fh-wels.at}\\ %
		\scriptsize University of Applied Sciences - Upper Austria%
		}
	\maketitle
Section~\ref{sec:ext_exp} contains extended experiments including results on our third dataset.
We also include the user evaluation  questionnaire as a separate file. The file contains links to videos which were seen by the participants. 
\input {./supplemental/supplementaryMaterial.tex}
\end{document}